---
title: "PML Prediction"
output: html_document
date: "2024-04-30"
---
**Introduction**

In this project, we used data gathered from accelerometers positioned on the belt, forearm, arm, and dumbbell of six participants. Our aim is to predict the manner in which they perform exercises. This projects falls within the study of the quantified self movement, where individuals utilize devices like Jawbone Up, Nike FuelBand, and Fitbit to collect extensive data on personal activity. Analyzing this data, enthusiasts seek to enhance their health, discern behavioral patterns, or simply indulge their tech-savvy inclinations. While quantifying the extent of physical activity is common, our focus extends to quantifying the quality of execution, a dimension often overlooked in such analyses.

```{r, echo = TRUE}
# Load libraries
library(caret)
library(randomForest)
library(ggplot2)

# Set seed for reproducibility
set.seed(12345)
```

**Process Data**

```{r, echo = TRUE}
# Function to preprocess data
preprocess_data <- function(data) {
    # Remove columns with high percentage of missing values
    threshold <- 0.6 * nrow(data)
    data <- data[, colSums(is.na(data)) < threshold]
    # Remove near zero variance predictors
    nzv <- nearZeroVar(data)
    data <- data[, -nzv]
    # Remove non-predictive variables
    data <- data[, !names(data) %in% c("X", "user_name", "raw_timestamp_part_1",
                                       "raw_timestamp_part_2", "cvtd_timestamp",
                                       "new_window", "num_window")]
    return(data)
}

# Load local data files
train <- read.csv("/Users/youknowjp/Downloads/pml-training.csv")
test <- read.csv("/Users/youknowjp/Downloads/pml-testing.csv")

# Apply preprocessing
train <- preprocess_data(train)
test <- preprocess_data(test)
```

```{r, echo = TRUE}
# Break the data into training and validation sets
trainIndex <- createDataPartition(train$classe, p = 0.75, list = FALSE)
training <- train[trainIndex, ]
validation <- train[-trainIndex, ]
```

**Use RandomForest to train data**

```{r, echo = TRUE}
# Train the random forest model
rfModel <- randomForest(as.factor(classe) ~ ., data = training, importance = TRUE, ntree = 50)

# Predict on the training set
ptraining <- predict(rfModel, training)

# Convert actual classes to factor if not already
training_classe <- factor(training$classe)
ptraining <- predict(rfModel, training, type = "class")  

# Ensure both factors have the same levels
all_classes <- levels(training_classe)

# Relevel predicted classes to include all possible classes from actual data
ptraining <- factor(ptraining, levels = all_classes)

# Now calculate the confusion matrix
confusionMatrixData <- confusionMatrix(ptraining, training_classe)
print(confusionMatrixData)

```

Cross validation accurracy is 100% & out-of-sample error is 0%. So our model performs good.

```{r, echo = TRUE}
# Predict on the test set
ptest <- predict(rfModel, test)
ptest
```
